### 2016-06-23 introduction.tex
edits

The guarantees provided by serializability, and strong isolation in
particular, do not come for free, however - pessimistic concurrency
control methods, for example, require databases to provide expensive
techniques such as two-phase locking that incur overhead to deal with
deadlocks, rollbacks, and re-execution~\cite{twopl,ullmanbook}.
Similar criticisms apply to optimistic multi-version concurrency
control methods that must deal with timestamp and version
management~\cite{BG81}.  Regardless of the particular concurrency
control method employed, enforcing serializability in a replicated
data store additionally requires global coordination among
geo-distributed replicas, which makes the store \emph{unavailable} in
the presence of network
partitions~\cite{cap,sernotavlbl,bailishat,bernsigmod13}.

The tension between serializable transactions, which are easy to
reason about but difficult to implement, and more pragmatic goals
driven by performance and availability considerations, has motivated
the development of different weaker forms of transaction isolation,
beginning as early as 1976~\cite{gray1976}. The ANSI SQL 92 standard
defines three such weak isolation levels which are now implemented in
many relational and NoSQL databases. Not surprisingly, weakly-isolated
transactions have been found to significantly outperform serializable
transactions on benchmark suites, both on single-node databases and
multi-node replicated stores~\cite{dbtuningbook,bailishat,bailisvldb},
leading to their overwhelming adoption. A 2013
study~\cite{bailishotos} of 18 popular ACID and ``NewSQL'' databases
found that only three of them offer serializability by default, and
half, including Oracle 11g, do not offer it at all.  A 2015
study~\cite{bailisferal} of a large corpus of database applications
finds no evidence that applications manifestly change the default
isolation level offered by the database. Taken together, these studies
make clear that weakly-isolated transactions are quite prevalent in
practice, and strongly-isolated (serializable) transactions are often
eschewed.

@@

Our results thus provide the first (to the best of our knowledge)
mechanism that precisely relates high-level program invariants to
low-level weak isolation and store consistency guarantees, thereby
allowing weakly-isolated transactions to enjoy the same rigorous
reasoning capabilities as their strongly isolated (serializable)
counterparts.

>>>>

The guarantees provided by serializability, and strong isolation in
particular, do not come for free, however - pessimistic concurrency
control methods that enforce serializability, for example, require
databases to provide expensive techniques such as two-phase locking
that incur overhead to deal with deadlocks, rollbacks, and
re-execution~\cite{twopl,ullmanbook}.  Similar criticisms apply to
optimistic multi-version concurrency control methods that must deal
with timestamp and version management~\cite{BG81}.  Regardless of the
particular concurrency control method employed, enforcing
serializability in a replicated data store additionally requires
global coordination among geo-distributed replicas, which makes the
store \emph{unavailable} in the presence of network
partitions~\cite{cap,sernotavlbl,bailishat,bernsigmod13}.

The tension between serializable transactions, which are easy to
reason about but difficult to implement, and more pragmatic goals but
weaker variants that are driven by performance and availability
considerations, has motivated the development of weaker forms of
transaction isolation, beginning as early as 1976~\cite{gray1976}. The
ANSI SQL 92 standard defines three such weak isolation levels which
are now implemented in many relational and NoSQL databases. Not
surprisingly, weakly-isolated transactions have been found to
significantly outperform serializable transactions on benchmark
suites, both on single-node databases and multi-node replicated
stores~\cite{dbtuningbook,bailishat,bailisvldb}, leading to their
overwhelming adoption. A 2013 study~\cite{bailishotos} of 18 popular
ACID and ``NewSQL'' databases found that only three of them offer
serializability by default, and half, including Oracle 11g, do not
offer it at all.  A 2015 study~\cite{bailisferal} of a large corpus of
database applications finds no evidence that applications manifestly
change the default isolation level offered by the database. Taken
together, these studies make clear that weakly-isolated transactions
are quite prevalent in practice, and strongly-isolated (serializable)
transactions are often eschewed.

@@

\noindent Our results thus provide the first (to the best of our knowledge)
mechanism that precisely relates high-level program invariants to
low-level weak isolation and store consistency guarantees, thereby
allowing weakly-isolated transactions to enjoy the same rigorous
reasoning capabilities as their strongly isolated (serializable)
counterparts.

The remainder of the paper is organized as follows.  The next section
provides motivation and background on serializable and weakly-isolated
transactions. \S\ref{sec:opsem} presents an operational semantics for
a core language that supports weakly-isolated transactions, and which
is parameterized over different isolation and store-consistency
notions. \S\ref{sec:proofrules} formalizes the proof systemt that we
use to reason about program invariants, and establishes the soundness
of these rules with respect to the semantics.  We describe the impact
of our reasoning framework in the context of several real-world case
studies in \S\ref{sec:casestudies}.  Related work and conclusions are
given in \S\ref{sec:relatedwork}.
